---
title: "Installing SpaDES"
author: "Alex M. Chubaty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing SpaDES}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

## Installing R

### Installing prerequisites

#### Windows

1. Download and run the R installer (as administrator!) from <https://cran.r-project.org/bin/windows/base>.

2. (Optional) To build and compile packages, download and install the latest Rtools (as administrator!) from <https://cran.r-project.org/bin/windows/Rtools/>.
    Be sure to check the "add rtools to PATH" option during installation.

3. Install GIS tools by downloading OSGeo4W from <http://download.osgeo.org/osgeo4w/osgeo4w-setup-x86_64.exe> and running the installer (select 'Express Install').

#### macOS

1. Download and run the R installer from <https://cran.r-project.org/bin/macosx>.

2. Install Xcode commandline tools from the terminal:

    ```
    xcode-select --install
    ```

3. Install homebrew from the terminal (more info at <https://brew.sh/>):

    ```
    /bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"
    ```

4. Install GDAL and other spatial packages:

    ```
    brew install pkg-config
    brew install gdal
    ```
    
    Then, install spatial R packages from source to pick up the new GDAL and PROJ libraries:
  
    ```{r install-spatial}
    ## (re)install spatial packages from source
    install.packages(c("rgeos", "sf", "sp"), type = "source")
    
    ## rgdal requires some additional configuration to build correctly:
    ##   based on http://stackoverflow.com/a/26836125/1380598
    install.packages("rgdal", type = "source",
                     configure.args = c("--with-proj-include=/usr/local/include",
                                        "--with-proj-lib=/usr/local/lib"))
    library(rgdal) ## confirm the GDAL version being used
    ```

5. Install OpenMP support for `data.table`:

    ```
    brew install llvm
    ```

6. Install GLPK (needed for full `igraph` functionality):

    ```
    brew install glpk
    ```

    Then, reinstall `igraph` from source so it is compiled with GLPK:

    ```{r install-igraph}
    install.packages("igraph", type = "source")
    ```

#### Linux

Install via `apt` or `yum` depending on your distribution.
Instructions for Ubuntu 20.04 LTS are shown below.

1. Install R and its development libraries:

    ```
    sudo apt install r-base r-base-dev
    ```

2. Install java and configure R to use it:

    ```
    sudo apt build-dep -y r-cran-rjava
    sudo R CMD javareconf
    ```

3. Install dependencies for spatial and other packages:

    ```
    sudo apt build-dep -y r-cran-tkrplot

    sudo apt-get install -y \
      curl \
      gdal-bin \
      git \
      libcairo2-dev \
      libcurl4-openssl-dev \
      libgdal-dev \
      libgeos-dev \
      libgit2-dev \
      libglpk-dev \
      libgmp-dev \
      libjq-dev \
      libmagick++-dev \
      libproj-dev \
      libprotobuf-dev \
      libprotoc-dev \
      libssh2-1-dev \
      libssl-dev \
      libudunits2-dev \
      libv8-dev \
      libxml2-dev \
      libxt-dev \
      pandoc \
      pandoc-citeproc \
      protobuf-compiler \
      python3-gdal \
      tcl-dev \
      tk-dev
    ```


### Installing Rstudio

**Recommended.**

Download and run the installer for your operating system from <https://www.rstudio.com/products/rstudio/download/#download>.
